---
title: "lab 3"
author: "Fin"
date: "9/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)

ggplot (data=cdc)+
  geom_histogram(mapping=aes(x=age),fill="grey")+
  geom_freqpoly(mapping=aes(x=age), color="black")+
  facet_grid(~smoke100)+
  theme_minimal()+
  labs(
    title="Effect of Smoking on Age"
  )
```
 
 
 The findings generated from this plot show that on average the age of smokers is centered around 50 with a large sharp drop off at around 60. This would indicate a shorter life expectancy for smokers. It can also be seen that a greater amount of young people do not smoke and it is may be a generational change that sees further decline with more time.
 
 
 For this project I created a plot comparing smoking and age and found that smoking is much more common among older people but is also a result that leads to earlier deaths.
 
 
 

