---
title: "lab 3 Feeling Dangerous"
author: "Fin,Nahom,Chun,Jake"
date: "9/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```
# Team Section

### Overall Question 
How does gender affect the health repsonses among non-smokers and smokers? (General health responses = excellent, very good, good, fair, and poor)

### Team Plot

```{r}
ggplot(data = cdc)+
  geom_bar(mapping = aes(x = smoke100, fill = genhlth), position = "dodge") + 
  facet_wrap(~gender)+
  labs(title = "Male Vs. Female Responses on Personal Health", 
       subtitle = "Non-Smokers vs. Smokers",
       caption = "0 = Non-Smoker | 1 = Smoker")
```

### Conclusion
From the two differnt plots, we have found that male smokers percieve themselves to have better health than female smokers. Interesting enough, we made the conclusion that **men** will manipulate or lie about their personal health more often than women. We can make this assumption due to the fact that **male smokers** respond with a higher frequency of "very good"" and "good" than male non-nonsmoker. **Female smokers** have an overall lower health response in every category. Our recomendation is for sampling to occur without requiring personal information such that people feel inclined to anwser although this problem will likely always exist. 



# Individual Section 

#### Jake

### Individual Question
How are general health responses affected by smoking cigarettes ? 

### Plots 
```{r}
ggplot(data = cdc) +
  geom_bar(mapping = aes(x = smoke100, fill = genhlth), position = "dodge") + 
  labs(title = "Smoker vs. Non-Smoker Responses on Personal Health", 
       caption = "0 = Non-Smoker | 1 = Smoker")
```

### Findings and Contributions
I, Jacob Boeckenstedt, created a plot showing how general health responses of non-smokers compare to the general health responses of smokers. It is not surprising to see that a significantly larger amount of respondents who have NOT smoked reported excellent health. However, it is surprising to see that respondents who smoke, report a higher count of "good" overall health. This lead me to believe that some people who smoke are either blinded by the effects or distort their helath for a snese of affirmation for themselves. I used the geom_bar function and adjusted the postiion and fill of the mapping to optimize the visual aspect and make it quick and easy to readIn addition I used a position adjustment to <<>> and changed the default labels to make the plot easier to read. In addition I had to use the mutate as. character function to make the x asis easy to read. This removed values like -.5, .75, 1.5 that were irrelevant. This contributed to the foundation of our group project where we then filtered and seperated male and females to see the differnces in responses there. 

----------------------------------

#### Fin Carlson

### Individual Question

What is the relationship between smoking and age?

### Plots 

```{r echo=FALSE, message=FALSE }
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)

```
 
```{r}
ggplot(data=cdc)+
  geom_histogram(mapping=aes(x=age),stat="bin", fill="gray",
  position = "stack",binwidth = 4.5, bins = 4.5)+
  geom_freqpoly(mapping=aes(x=age), color="black",binwidth=5, bins=5)+
  facet_grid(~smoke100)+
  theme_minimal()+
  labs(
    title="Distribution of Age for Smokers and Non-smokers"
  )+
  ylab("Amount of People")+
  xlab("Age")
```
 
### Findings

 The key finding generated by this plot is that those who smoke seem to have lower life expectancy. Looking at the plot for smokers, a sharp drop off is seen at around seventy years of age. This drop off contiunes as seen in the frequency polygon plot and in general the line is much steeper for the smokers after sixty five years of age. This finding could be disputed by the thought that people are simply more likely to quit at around this age however this seems unlikely and the sheer size of the drop off indicates this is not true. 
 The plot also suggests smoking may be a generational pastime that will die out with future generations. The plot for smokers is centered around fifty years of age with the data being skewed left. More people older than fifty smoke than people younger than fifty. The oppoisite is true for the non-smokers plot as the data is centered much closer to thirty five years of age and skewed right, with more people younger than thirty five not smoking than people older. This may be disputed by the legal age of smoking being eighteen years of age however the data shows a small jump for smokers at around this age however it is minamal and the overall picture still suggests a generational gap. This is also supported by new smoking technologies and policies.

### Contributions 

  For this lab assignment I created a plot that explored the relationship between age and whether a person was a smoker. I started by creating a simple histogram that I then wrapped to show the differences between those who smoke and those who do not. I wanted to better see the overall change in my distribution and added a frequency polygon plot to better visualize this change. I was presented with an error in the default size of my binwidth and as result had to manually set for both plots. I also looked through the theme() function and decided on minimal for my plots. The final task I completed was organzing all of our work into a single Rmarkdown and making sure it could knit before turning it in.

----------------------------------------------------------------

#### Nahom

### Individual Question

What is the percentage that makes up the smoker and non smoker population of each gender?


```{r echo=FALSE, message=FALSE}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```

### Plots

```{r}
ggplot(data =cdc) +
geom_bar(mapping = aes(x = smoke100, fill = gender), position = "fill") + 
  facet_grid(~smoke100) + ggtitle("Percentage of Smokers and Non-smokers between Male and Female")
```

### Findings

Based on the bar graph, of non-smokers population Females make up about 60% of it while Males make up the rest 40%, Of the smokers Males roughly makeup 49% of the population and Females makeup the other 51%. 

### Contributions




-----------------------------------


#### Chun

### Individual Question

Do smokers usually have a health plan?

### Plots  

```{r, }
ggplot(data=cdc)+
  geom_bar(mapping=aes(x=smoke100,fill=hlthplan))+
  facet_grid(~hlthplan)
  
```

### Findings

According to the bar chart, most smokers have a health plan. Also, for those who have a health plan，less people have smoked over 100 cigarettes.

### Contributions
 
 

