---
title: "lab 3 Feeling Dangerous"
author: "Fin,Nahom,Chun,Jake"
date: "9/12/2019"
output: html_document
---
# Team Section

### Overall Question 

### Team Plot

### Conclusion


# Individual Section 

#### Jake

### Individual Question

### Plots 

### Findings 

### Contributions

----------------------------------

#### Fin Carlson

### Individual Question

What is the relationship between smoking and age?

### Plots 

```{r echo=FALSE, message=FALSE }
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)

```
 
```{r}
ggplot(data=cdc)+
  geom_histogram(mapping=aes(x=age),stat="bin", fill="gray",
  position = "stack",binwidth = 4.5, bins = 4.5)+
  geom_freqpoly(mapping=aes(x=age), color="black",binwidth=5, bins=5)+
  facet_grid(~smoke100)+
  theme_minimal()+
  labs(
    title="Distribution of Age for Smokers and Non-smokers"
  )+
  ylab("Amount of People")+
  xlab("Age")
```
 
### Findings

 The key finding generated by this plot is that those who smoke seem to have lower life expectancy. Looking at the plot for smokers, a sharp drop off is seen at around seventy years of age. This drop off contiunes as seen in the frequency polygon plot and in general the line is much steeper for the smokers after sixty five years of age. This finding could be disputed by the thought that people are simply more likely to quit at around this age however this seems unlikely and the sheer size of the drop off indicates this is not true. 
 The plot also suggests smoking may be a generational pastime that will die out with future generations. The plot for smokers is centered around fifty years of age with the data being skewed left. More people older than fifty smoke than people younger than fifty. The oppoisite is true for the non-smokers plot as the data is centered much closer to thirty five years of age and skewed right, with more people younger than thirty five not smoking than people older. This may be disputed by the legal age of smoking being eighteen years of age however the data shows a small jump for smokers at around this age however it is minamal and the overall picture still suggests a generational gap. This is also supported by new smoking technologies and policies.

### Contributions 

  For this lab assignment I created a plot that explored the relationship between age and whether a person was a smoker. I started by creating a simple histogram that I then wrapped to show the differences between those who smoke and those who do not. I wanted to better see the overall change in my distribution and added a frequency polygon plot to better visualize this change. I was presented with an error in the default size of my binwidth and as result had to manually set for both plots. I also looked through the theme() function and decided on minimal for my plots.

----------------------------------------------------------------

#### Nahom

### Individual Question

What is the percentage that makes up the smoker and non smoker population of each gender?


```{r echo=FALSE, message=FALSE}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```

### Plots

```{r}
ggplot(data =cdc) +
geom_bar(mapping = aes(x = smoke100, fill = gender), position = "fill") + 
  facet_grid(~smoke100) + ggtitle("Percentage of Smokers and Non-smokers between Male and Female")
```

### Findings

Based on the bar graph, of non-smokers population Females make up about 60% of it while Males make up the rest 40%, Of the smokers Males roughly makeup 49% of the population and Females makeup the other 51%. 

### Contributions




-----------------------------------


#### Chun

### Individual Question

Do smokers usually have a health plan?

### Plots  

```{r, }
ggplot(data=cdc)+
  geom_bar(mapping=aes(x=smoke100,fill=hlthplan))+
  facet_grid(~hlthplan)
  
```

### Findings

According to the bar chart, most smokers have a health plan. Also, for those who have a health plan，less people have smoked over 100 cigarettes.

### Contributions
 
 

