---
title: "lab 3"
author: "Nahom Taye"
date: "9/11/2019"
output: html_document
---

#### Overall Qusetion

What is the percentage that makes up the smoker and non smoker population of each gender?

#### Answer to question

Based on the bar graph, of non-smokers population Females make up about 60% of it while Males make up the rest 40%, Of the smokers Males roughly makeup 49% of the population and Females makeup the other 51%. 

#### Team Plot
```{r}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```

#### Recommendation

Recommendation to CDC
#### Series of Plots and Findings

```{r, echo=FALSE, message=FALSE}
ggplot(data =cdc) +
geom_bar(mapping = aes(x = smoke100, fill = gender), position = "fill") + 
  facet_grid(~smoke100) + ggtitle("Percentage of Smokers and Non-smokers between Male and Female")
```


###Do smokers usually have a health plan?
```{r, echo=FALSE, message=FALSE}
ggplot(data=cdc)+
  geom_bar(mapping=aes(x=smoke100,fill=hlthplan))+
  facet_grid(~hlthplan)
  
```
###According to the bar chart, most smokers have a health plan. Also, for those who have a health plan，less people smokes over 100 cigarettes.
