---
title: "lab 10 Feeling Dangerous"
author: "Fin Carlson,Nahom,Chun,Jake Boeck"
date: "11/1/2019"
output: html_document
---

```{r setup, include=FALSE}
#Loading Data
library(tidyverse)
library(ggplot2)
answers <- read_csv("Answers_trunc.csv")
questions <- read_csv("Questions_trunc.csv")

merged <- answers %>%
  left_join(questions, c('ParentId' = 'Id')) 

merged <- merged %>% 
  mutate(total_time = CreationDate.x - CreationDate.y) %>%
  filter(total_time>0)

```
# Team Section

### Team Plots

```{r, echo=FALSE}
#timeliness of score
ggplot(data = (merged)) + 
  geom_point(mapping = aes(x = total_time, y = Score.x), size=0.1)+
   xlim(c(0, 3000))+
   ylim(c(0, 3000))+
  labs(title = "Time taken to Answer vs. Score of post", 
       x = "Time", y = "Score")


#normal distribution?
ggplot(merged) + geom_histogram(aes(Score.x), bins=100) +
     xlim(c(0, 100))+
labs(title = "Checking if score is Normally Distributed ", 
       x = "Score", y = "Count")

```

### Conclusions

When comparing the timeliness of a post against it’s score, it can be noted that in general there is little relationship between the time passed and the score. It should be realized however that the highest scores very rarely occur after a total time of 1000 minutes.

When considering using Z-scores to compare different scores, it should be noted that the distribution does not appear normal and as a result z-scores should not be used for this data set.

Looking over the results of our individual functions it should be realized that they are not super accurate. This is likely due to the fact they focus on detection of one word or short phrases. The data presented varies greatly in the word usage as well as the score distribution and as such these metrics are not accurate in predicting score. The most accurate of the functions can be found under Fin Carlson's section (graph 1) and it uses the small phrase "How do I". This is likely more accurate because it depends on a phrase instead of a single word and because it identifies a question as being somewhat specific. This makes it easier to respond to and likely causes higher scores as responders know how to go about answering. In general the ideal question can be identified as being more specific and therefore simpler to answer.

# Individual Section 

#### Jake

### Individual Features

### Contributions

----------------------------------

#### Fin Carlson

### Individual Features

##### Proportions of High Scores for Q&A Features

```{r, echo=FALSE}
#Question Function
q_tried <- questions %>%
  mutate(tried = str_detect(Title, "How do I")) %>%
  mutate(s = case_when((Score>=41)~"high", ((Score<41)~"low")))

q1_tried <- q_tried %>% 
  mutate(s.num = ifelse(Score>=41, 1, 0))

q1_tried %>% group_by(tried) %>% summarise(mean(s.num))


#Anwser Function
a_you <- answers %>%
  mutate(you = str_detect(Body, "you")) %>%
  mutate(s = case_when((Score>=20)~"high", ((Score<20)~"low")))

a1_you <- a_you %>% 
  mutate(s.num = ifelse(Score>=41, 1, 0))

a1_you %>% group_by(you) %>% summarise(mean(s.num))
```

I designed my first feature to search the Title of the Questions data for any instance of "How do I". This would indicate that the question is rather specific and as a result can be more easily answered by a responder. I chose this feature because I often ask questions that are direct and reference a specific topic and I wanted to see if this was in general more effective. I believe there is a positive relationship between score and including this phrase as a higher proportion of those that included the phrase had what was considered a high score (greater than 40). This would make sense as those responding can easily understand the problem.

I designed my second feature to search the body of the Answers data for any instance of "you". I chose this word because it would indicate the answer is directly responding to the original poster which would likely make the answer better. After mutating the data however I do not believe there is a relationship between my feature and overall score as there are very similar proportions of high scores for answers that include and don't include "you". 

### Plots

```{r, echo=FALSE}
ggplot(data = q_tried) +
  geom_bar(mapping = aes(s, fill = s)) +
  facet_wrap(~ tried) +
  labs(title = "Effect of including HOW DO I on score for Questions", 
       x = "Score", y = "Count",
       fill="Score")

ggplot(data = a_you) +
  geom_bar(mapping = aes(s, fill = s)) +
  facet_wrap(~ you) +
  labs(title = "Effect of including YOU on Score for Answers", 
       x = "Score", y = "Count",
       fill="Score")
```


### Contributions 

For this lab I created my two features as well as our team plots to answer the time and normality questions. I provided conclusions for our team section and compiled all of our work into one document that I knitted and turned in. After creating our final markdown it became apparent that my teammates used the same graphs as me even though my final product was on our shared github long before anyone else and I had communicated this fact. As a result I would hope my grades are not affected by this as it was out of my control and can be cited to problems already brought to the Professor's attention.

----------------------------------------------------------------

#### Nahom

### Individual Features

```{r, echo = FALSE}
Qmean<- mean(questions$Score)
Amean <- mean(answers$Score)
question_Python <- questions %>%
  mutate(Python = str_detect(Title, "Python")) %>%
  mutate(result_Q = case_when((Score>= Qmean)~"high", ((Score< Qmean)~"low")))
answer_Python <- answers %>% 
  mutate (Python = str_detect(Body, "Python")) %>%
  mutate (result_a = case_when((Score >= Amean)~"HIGH",((Score< Amean)~"low")))
```

My feature for the Question is "Python" I think this is important becauses it helps better understand how much more or less difficult Questions involving or questions about Python compared to other questions , which it is based on my first graph, and knowing this can help people who are answering the questionnaire can be better prepared for alike questions to get a better result.

### Plots

```{r, echo = FALSE}
ggplot(data = question_Python) +
  geom_bar(mapping = aes(result_Q, fill = result_Q)) +
  facet_wrap(~ Python) +
  labs(title = "Scores of Questions involving Python", 
  x = "Scores", y = "Count") + scale_y_continuous(c(0, 6000))
ggplot(data = answer_Python) +
  geom_bar(mapping = aes(result_a, fill = result_a)) +
  facet_wrap(~ Python) +
  labs(title = "Scores of Answers involving Python", 
  x = "Scores", y = "Count") + scale_y_continuous(c(0, 4000))
```


-----------------------------------


#### Chun

### Individual Features

```{r, echo=FALSE}
Q = questions %>% mutate(feature = str_detect(Title, "What"))
m1=mean(Q$Score, na.rm = FALSE)
Q=Q%>%mutate(score = case_when((Score<m1)~"Low", ((Score>m1)~"High")))

A = answers %>% mutate(feature = str_detect(Body, "This"))
m2=mean(A$Score, na.rm = FALSE)
A=A%>%mutate(score = case_when((Score<m2)~"Low", ((Score>m2)~"High")))

Qm=Q%>%group_by(score)%>%summarise(mean(feature))
Am=A%>%group_by(score)%>%summarise(mean(feature))
```


My feature for the question is "What" and I think it is rather important because usually, when people using "what" in the question, there are really don't understand the idea, definition or meaning of the problem or function they might use when coding.(And of course, this is the word I always use when I have a coding problem.) And I choose "This" as the feature for the answers since it is one of the most common way to answer the question starting with "What". I think question with these features usually earn a higher score since these are the basic knowledge for anyone who learn to code and these Q & A can save a lot of people.

### Plots

```{r, echo=FALSE}
ggplot(data = A) +
  geom_bar(mapping = aes(score, fill = score)) +
  facet_wrap(~ feature)

ggplot(data = Q) +
  geom_bar(mapping = aes(score, fill = score)) +
  facet_wrap(~ feature)

```

 
 