---
title: "lab 14 Feeling Dangerous"
author: "Fin,Nahom,Chun,Jake"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)
library(modelr)

#loading and tidying data
data <- read.csv("nba_2017_nba_players_with_salary.csv")
data <- data %>% select(2:39)
data1 <- data %>% rename_at(vars(starts_with("X")), 
                                        funs(str_remove(., "X")))
data1 <-as_tibble(data1)
```
# Team Section

### Overall Question and Its Importance

How does an increase in salary affect the the minutes played per game for players who play a substainal amount of an NBA game?(over 21 mintues)

Our teams objective is to determine the most overvalued and undervalued players for the 2017 NBA Regular Season.


### Dataset Description

The dataset being anaylzed contains the 2017-2018 season averages for each player in the NBA. 

### Team Findings
```{r include=FALSE}
#FIN 
#EDA for Games Played vs. salary 
ggplot(data=data1)+
  geom_point(mapping=aes(x=GP, y=SALARY_MILLIONS))+
  geom_smooth(mapping=aes(x=GP, y=SALARY_MILLIONS))
  
  ##data is to clumped to make assumptions, GP not a good factor
  summary(data1$GP)
```

```{r, include=FALSE}
#FIN
#EDA for minutes per game vs. salary
ggplot(data=data1)+
  geom_point(mapping=aes(x=MPG, y=SALARY_MILLIONS))+
  geom_smooth(mapping=aes(x=MPG, y=SALARY_MILLIONS))

#code for anaylzing value based on mpg  
  
  ##avg salary 7.3
  ##avg mpg 21.5
  ##avg wins 2.86
  ##med wins 1.4
summary(data1$MPG)
summary(data1$SALARY_MILLIONS)
summary(data1$WINS_RPM)

  ##avg salary= 10.5 if mpg>21
min_above <-data1 %>% filter(MPG>= 21.57)

data1 %>% filter(MPG>= 21.57) %>% summarise(mean_data=mean(SALARY_MILLIONS))

  ##creates group of players who play more than average and are paid less than average
minutesSalary <-data1 %>% filter(MPG>= 21.57 & SALARY_MILLIONS<=10.5) 


  
  ##visualization of group
ggplot(data=minutesSalary)+
  geom_point(mapping=aes(x=MPG, y=SALARY_MILLIONS))+
  geom_smooth(mapping=aes(x=MPG, y=SALARY_MILLIONS))

  ##further filtering based on graph results

summary(minutesSalary$MPG) #avg mpg for this group is 27.8
summary(minutesSalary$SALARY_MILLIONS) #avg salary for this group is 4.85

#creating even smaller group based on group average
Finalminutes1<- minutesSalary %>% 
  filter(MPG>= 27.78 & SALARY_MILLIONS<=4.851) %>% arrange(MPG) 


##filters out truly worse than avg players using group avg of holistic stat 
Finalminutes2<- Finalminutes1 %>% filter(WINS_RPM>=2.86) 


##seeing who makes the most, and who is most benefitial 
leastPaid <- Finalminutes2 %>% 
  select(PLAYER, POSITION, MPG, SALARY_MILLIONS, WINS_RPM) %>% 
  arrange(SALARY_MILLIONS)
leastPaid

Best <- leastPaid %>% arrange(desc(WINS_RPM))
Best


## Further filtering undervalued group to show the top 3
min_above %>% filter(WINS_RPM>10)  #10/342 .03 percent of players have winshare>10 
Finalminutes3<- leastPaid %>% filter(WINS_RPM>10) 
Finalminutes3

```
### Team Plots
```{r}
regresh0 <- lm(MPG ~ SALARY_MILLIONS, data = pos)
coef(regresh0)

ggplot(min_above,aes(x=SALARY_MILLIONS,y=MPG))+
  geom_point() +
  geom_smooth(se=F, color="grey") +
  geom_point(Finalminutes3, mapping=aes(x=SALARY_MILLIONS,y=MPG, color=PLAYER))+
  geom_abline(intercept = regresh0$coefficients[1], slope = regresh0$coefficients[2], colour = "red", size = .8)+
labs(title = "Salary vs. MPG for the 2017 NBA Season",
       subtitle = "Most undervalued players are shown in color",
       caption = "Source: NBA.com", 
       x = "Salary (Millions)", y = "Minutes per Game")
```
### Conclusion
After applying a linear model to Minutes per Game and Salary we found that Nikola Jokic falls below the line of best fit. The linear regression line over predicts Jokic’s salary while the non linear fitted line under predicts his salary. From our linear regression, we can interpret the relationship between salary and minutes as an increase in salary by a million dollars will increase the minutes played by about 12 seconds or .2 mintues.

#### Jake

### Individual Question
How does minutes played affect points per game? Break it down by position: what is the difference between the smallest and tallest players (centers and point gaurds)?

How does age affect minutes played for centers? Since they have a lot bigger bodies, they tend to get tired quicker and generally retire before other positions in the NBA.

### Tools I Used
I used the lm function to get the coefficents and residuals for my linear regression. I onnly ran two variables against each other and I did not plot a graoh on the residuals but just the different variables.

### Plots
```{r, echo=FALSE}
point10 <- data1 %>% filter(POINTS >= 10)
pos <- point10 %>% filter(POSITION=="C"  | POSITION == "PG" | POSITION == "SF")
center <- point10 %>% filter(POSITION=="C")
pg <- point10 %>% filter(POSITION=="PG")
sf <- point10 %>% filter(POSITION=="SF")
Tseven <- data1 %>% filter(POINTS >=27)

# lm for the coefficents and residuals
regresh <- lm(POINTS ~ MP, data = pos)
coef(regresh)

centers <- lm(POINTS ~ MP, data = center)
coef(centers)

pgs <- lm(POINTS ~ MP, data = pg)
coef(pgs)

sfs <- lm(POINTS ~ MP, data = sf)
coef(regresh)

points <- ggplot(data=pos)+
  geom_point(mapping=aes(x=MP, y=POINTS, colour=POSITION))+
  geom_abline(intercept = regresh$coefficients[1], slope = regresh$coefficients[2], colour = "yellow", size = 2) +
  geom_smooth(se=F,mapping=aes(x=MP, y=POINTS, colour=POSITION))
print(points + labs(title="Points vs Minutes Played (averaging 10PPG+)",
                    y="Points per game", x = "Minutes Played"))
```
### Findings 
You can interpret this linear regression anlaysis such that; if there is a one minute increase in time played per game, you can expect an .94 more points per game. For Centers you can expect points pergame to increase .75 points for every minute increase. And for point gaurds you can expect points to increase 1.23 for every minute.

```{r, echo=FALSE}
regresh2 <- lm(AGE ~ MP, data = center)
coef(regresh2)

age <- ggplot(data=center)+
  geom_point(mapping=aes(x=MP, y=AGE, colour=POSITION))+
  geom_smooth(mapping=aes(x=MP, y=AGE, colour=POSITION))+
  geom_abline(intercept = regresh2$coefficients[1], slope = regresh2$coefficients[2], colour = "green", size = .8)
print(age + labs(title="Age vs. Minutes Players (averaging 10PPG+)",
                    y="Age", x = "Minutes Played"))
```
### Findings 
You can interpret this linear regression anlaysis such that; with a one minute increase in minutes played, you can expect an increase in age by about a month for the average center in the NBA.

### Contributions
For this lab I went back into one of our previous labs and ran linear regressions on different data sets. I completed our team plot and completed the findings for our team. This included my interpretaion of the linear regression model I ran.

----------------------------------

#### Fin Carlson

### Individual Question

### Tools I Used

### Plots 

```{r}

```
 
### Findings

### Contributions 

----------------------------------------------------------------

#### Nahom

### Individual Question

### Tools I Used

### Plots

```{r}

```

### Findings


### Contributions


-----------------------------------

#### Chun

### Individual Question

### Tools I Used

### Plots  

```{r}

```

### Findings


### Contributions
 
