---
title: "lab 15 Feeling Dangerous"
author: "Fin Carlson,Nahom,Chun,Jake"
date: "12/13/2019"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(stringr)
library(modelr)

data <- read_excel("/Users/fin/Desktop/Data_Science/Fourteen_Data.xls")

View(data)

```

# Team Section

### Domain Expert, Overall Question and Dataset Description

### Team Findings and Recommendation

### Team Plots

-------------------------------------------

#### Fin

### Individual Question

### Tools I Used

### Findings, Recommendations, and Implications

### Plots

-------------------------------------------

#### Jake

### Individual Question

### Tools I Used

### Findings, Recommendations, and Implications

### Plots

-------------------------------------------

#### Chun

### Individual Question

### Tools I Used

### Findings, Recommendations, and Implications

### Plots

-------------------------------------------

#### Nahom
The domian expert is a PM at Colorado Department of Natural Resources that is asking if the elvation of the peaks affect the traffic?


### Individual Question
Is there a direct correlation between Elevation of a peak and its traffic? 


### Tools I Used
I used lm function, filter and mutate functions as well as ggplot functions.  

### Findings, Recommendations, and Implications

Theres is a weak correaltion between the elevation and the tarffic as seen on the graph. 

### Plots
```{r}
library(tidyverse)
peaks <- read.csv("/Users/olio/Downloads/Data Science 2019/Lab 15/fourteener.csv")
head(peaks)
```

```{r }
peaks %>% filter(fourteener == "Y") %>% mutate(Avg.Traffic = (Traffic.High + Traffic.Low)/2)

Avg.Peak <- mutate(peaks, Avg.Traffic = (Traffic.High + Traffic.Low)/2)
```

```{r }
ggplot(data=peaks, aes(x= Elevation_ft , y= Traffic.Low )) +
  geom_line()+
  geom_point()

ggplot(data=peaks, aes(x= Elevation_ft , y= Traffic.High)) +
  geom_line()+
  geom_point()

Avg.Lm <-lm(Elevation_ft ~ Avg.Traffic, data =Avg.Peak)
coef(Avg.Lm) ##Intercept =1.412605e+04
            ##Avg.Traffic = 4.183450e-03
regresh <- lm(Elevation_ft ~ Avg.Traffic, data = Avg.Peak)
coef(regresh)

ggplot(data=Avg.Peak, aes(x= Elevation_ft , y= Avg.Traffic)) +
  geom_line()+
  geom_point() + 
  geom_abline(intercept = regresh$coef[1], slope = regresh$coef[2], colours(red))+
  ggtitle(label = "Correlation between Peak elevation and Average Traffic")

par(mfcol = c(2,2))

```

-------------------------------------------

# Reflections 

#### Team

#### Fin

#### Jake

#### Chun

#### Nahom
My goals have not changed as my goals requrire me to graduate from college. In this course I have gained a good understanding of R and new statistical functions. One thing I could've done more reading the course book as it contains alot of usefull examples for R.  

-------------------------------------------

# Contributions 

#### Fin

#### Jake

#### Chun

#### Nahom
