---
title: "lab 13 Feeling Dangerous"
author: "Fin Carlson,Nahom,Chun,Jake Boeck"
date: "11/20/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)


```
# Team Section

### Overall Question

### Tests, Plots, and Conclusions

```{r, echo=FALSE}

```

--------------------------------------------------------------

# Individual Section 

## Jake

```{r setup, include=FALSE}
library(tidyverse)
library(gapminder)

inflation <- read_csv("inflation_annual_percent.csv")
gni <- read_csv("gnipercapita.csv")
```

### Overall Question

I chose to investigate the data sets under the economy tab, specifically, I chose the data set that measures inflation by annual percentge. I also choose to investigate the data set that measures gross national income per capita. I didn't realize that we had to investigate a data set on energy until after I finished but my major is economics so I found that to be the most interesting of all the data sets. 

The data sets include anual inflation percentge for all of the countries in this world as well as gross national income per capita for all of the countries. I preprocessed it by filter the data from years 1990 and on and only looked at the United States and United Kindom because the currencies are comprable and highly valued.  

```{r, include=FALSE}

inflation1 <- inflation %>% filter(country=="United States"|country=="United Kingdom") 
gni1 <- gni %>% filter(country=="United States"|country=="United Kingdom")

inflation1 <- inflation1 %>% gather(key = "Year", value = "inflation", -country) %>% filter(Year>=1990)

USinflation <- inflation %>% filter(country=="United States") %>% 
  gather(key = "Year", value = "inflation", -country) %>%
  filter(Year>=1990)

USgni <- gni %>% filter(country=="United States") %>% 
  gather(key = "Year", value = "gnipercapita", -country)

UKinflation <- inflation %>% filter(country=="United Kingdom") %>% 
  gather(key = "Year", value = "inflation", -country)

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms"
  perm_cor_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    y_shuffled <- sample(y)
    x_shuffled <- sample(x)
    
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    perm_cor_diffs[i] <- cor(y_shuffled, x_shuffled)
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(perm_cor_diffs)
}

USinflation_vector <- USinflation$inflation
USgni_vector <- USgni$gnipercapita

real_corr <- cor(USinflation_vector, USgni_vector) # -.404
perm_corr <- perm_cor(perms = 1000, USinflation_vector, USgni_vector)
perm_corr_tibble <- as_tibble(perm_corr)

```

### Tests, Plots, and Conclusions
I conducted a two sample permutation test using correlation as my test statistic. My null hypothesis: the correlation is equal to 0. Alternative hypothesis: the correlation was not equal to 0. I found the correlation between inflation and gross national income per capita is -.404 which fell in the 100th percentile for the permutation data. This means the null hypotheis can be rejected as we can be confident that caluculated correlation was not found by random chance alone. A histogram showing the results of the permutation is given below.

I also created a graph exploring the differences in anual inflation between the US and the UK. I created a geom point for this exploration. I found a strong postive correlation between inflation rates for each country. The inflation rates folow similar trends.

```{r}
# Plot Code
p <- ggplot(data = perm_corr_tibble) +
  geom_histogram(mapping= aes(x=value, col='red'), bins=100) +
  labs(title = "Permutaion Test for Correlation Between U.S. Inflation % and U.S. GNI") +
  xlab("Correlation")
p + scale_fill_brewer(direction = 0) + theme_dark() +
  theme(legend.position = "none")

ggplot(inflation1, aes(Year, inflation))+
  geom_point(aes(colour = country)) +
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Anual Inflation % for United States and United Kingdom") +
  xlab("Year")
```

### Map Function

Below is my map function. I created a vector with infaltion for each country United States and United Kingdom. My map function calculates the average anual infaltion percentage.
```{r, warning=FALSE}
map_vector <- inflation[inflation1$country == "United States" |
                         inflation1$country == "United Kingdom",]

map_vecotr <- map(map_vector, mean)
```
### Contributions

For this lab I met up with Fin to dicuss the datasets and questions we wanted to investigate. Fin was able to walk me through the permutaion tests he created. I was only familiar with t tests and prop tests so Fin was of great help for me on this lab. From there I completed my section with Fin and dicusseded team plots with him.

----------------------------------

## Fin Carlson

```{r Fin_setup, include=FALSE}
#loading in agriculture data
agg_data <- read_csv("agricultural_land_percent_of_land_area.csv")
#gathering agriculture data into 3 variables
agg_data1 <- agg_data %>% gather(key = "Year", value = "%_agg_land ", -country)

#loading in forrest data
for_data <- read_csv("forest_coverage_percent.csv")
#gathering forrest data into 3 variables
for_data1 <- for_data %>% gather(key = "Year", value = "%_for_land ", -country)

#joining data sets and filitering years with no data
data <- left_join(agg_data1, for_data1, by = c("Year", "country"))
data <- data %>% filter(Year >=1990)
```

### Overall Question

For this lab, I looked at data sets concering % of territory covered in farm and forrest land. I thought it would it be interesting to see how these two variable have changed over the years around the world. I preproccessed the data by first using the gather function on the two data sets to properly represent year as a single variable instead of indivual columns. I then joined the two data set's together using a left_join and I filtered out all years before 1990 as there was no data before this for the forestry set. I focused my main question on determining if a correlation existed between % farm land and % forrest land. 

### Tests, Plots, and Conclusions
```{r perm_test, include=FALSE}

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms"
  perm_cor_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    y_shuffled <- sample(y)
    x_shuffled <- sample(x)
    
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    perm_cor_diffs[i] <- cor(y_shuffled, x_shuffled, use="complete.obs")
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(perm_cor_diffs)
}

#Creating vectors for Forrest and Agriculutre data to be used on permutation function
for_vector <- data$`%_for_land `  
agg_vector <- data$`%_agg_land `

#Finding the real correlation between the two data sets (use is instead of na.rm)
cor_real<- cor(for_vector, agg_vector, use="complete.obs") #-.466

#running permutation function 
cor_vector <- perm_cor(perms = 1000, for_vector, agg_vector) 

#Changing permutation results from vector to tibble for graphing
cor_graph <- as_tibble(cor_vector)

#calculating percentile
(sum(cor_real<cor_vector)/1000)*100

```

I conducted a two sample permutation test using correlation as my test statistic. My null hypothesis stated that the correlation was equal to 0 while my alternative hypothesis stated that correlation was not equal to 0. I found the correlation between % farm and % forest land to be -.45 which fell in the 100th percentile for the permutation data. This means the null hypotheis can be rejected as we can be confident that caluculated correlation was not found by random chance alone. A histogram showing the results of the permutation is given below.

```{r, echo=FALSE}
ggplot(data = cor_graph)+
  geom_histogram(mapping= aes(x=value))+
  labs(title = "Permutation Test for Correlation between % Farm land and % Forest land")+
  xlab("Correlation")
   
  
ggplot(data = cor_graph)+
  geom_histogram(mapping= aes(x=value))+
  geom_vline(xintercept = -.466, color="red") +
  labs(title = "Permutation Test for Correlation between % Farm land and % Forest land",
       subtitle = "                                     Red line represents actual correlation")+
  xlab("Correlation")
```

### Map Function

```{r, warning=FALSE}
map_vector <- for_data[for_data$country == 'United States' |
                         for_data$country == 'Mexico' |
                         for_data$country == 'Canada', ]

map_vector <- map(map_vector, mean)
```

Given above is my map function. I first created a vector with the forest data for all countries in North America. My map function then calculates the average % of forest land for all of North America for every year. I only used this function in my exploratory anaylsis.

### Contributions

For this lab I completed my indivual section and helped complete the team section. I then  created the final document for our team and organized our work before turning it in.

----------------------------------

## Nahom

### Overall Question
How does the usage of electricity and coal usage per person compare to each other as the world is turning to clean enery
living, to answer this question I first gathered my data sets as year and country variable and coal and electricity usage per
person and cleaned it up even more by string dropping the 'X's in the year coloumn. 

### Tests, Plots, and Conclusions
```{r}
electricity <- read.csv(file="/Users/olio/Downloads/Data Science 2019/lab 13/electricity_use_per_person.csv")
coal <- read.csv(file="coal_consumption_per_cap (1).csv")


electricity_tidy <- electricity %>% gather(key="Year", value = "EPP", -country )
electricity_tidy %>% mutate(Year = str_remove(Year, 'X'))
coal_tidy <- coal %>% gather(key="Year", value = "CPP", -country)
coal_tidy %>% mutate(Year = str_remove(Year, 'X'))

joined <- left_join(electricity_tidy, coal_tidy, by =c("Year","country"))

```
```{r perm_test}
perm_cor <- function(perms= 1000, x,y) 

  {
  perm_cor_diffs <- numeric(perms)
  for(i in c(1:perms))
    {
    y_rand <- sample(y)
    x_rand <- sample(x)
    
    perm_cor_diffs[i] <-  cor(y_rand, x_rand, use = "complete.obs") #== 0.3
  }
  return(perm_cor_diffs)
}
electricity_vector <- joined$ `EPP`
coal_vector <- joined$ `CPP`

cor_real<- cor(coal_vector, electricity_vector)
cor_vector <- perm_cor(perms= 1000, electricity_vector, coal_vector)
cor_data <- as.tibble(cor_vector)
```
```{r}

ggplot(cor_data)+
  geom_density(mapping = aes(x=value))+
  labs(title = "Premutation Test for correlation between Coal usage and Electricity usage per person")+
  xlabs("Correlation")
```
### Map Function
```{r}
map_vector <- electricity[electricity$country =="China",
                        electricity$country =="United States",]
map_vector <- Map(map_vector, mean)

```

### Contributions
I completed my indiviual section and turing it in on time.

----------------------------------

###Chun Ni

```{r setup, include=FALSE}
library(tidyverse)
library(gapminder)
arms.o = read_csv("arms_imports_us_inflation_adjusted.csv")
oil.o = read_csv("oil_production_total.csv")
arms = arms.o %>% gather(key="Year", value = "AI", -country )

oil = oil.o %>% gather(key="Year", value = "OP", -country)
my.data = left_join(arms, oil, by = c("Year", "country"))
my.data = my.data %>% filter(Year >=1990)
my.data = na.omit(my.data)
```
### Overall Question
War always start with reasons. Resource conflict is often involved in it. In this lab, I want to know is oil really involved in it. To see the relation between them, I pick the oil production total as the data which represent the oil and weapon import as the possible area where war is going to happen. 

### Tests
Null hypothesis: the correlation between oil production and weapon import is equal to 0. 
Alternative hypothesis: the correlation between oil production and weapon import is not equal to 0.
```{r perm_test}
perm.cor <- function(perms= 1000, x,y) 

  {
  perm.cor.diffs <- numeric(perms)
  for(i in c(1:perms))
    {
    r.x = sample(x)
    r.y = sample(y)
    
    perm.cor.diffs[i] =  cor(r.x, r.y)
  }
  return(perm.cor.diffs)
}
arms.vector = my.data$ `AI`
oil.vector = my.data$ `OP`

cor.r = cor(arms.vector, oil.vector) #0.29
cor.p = perm.cor(perms = 1000, arms.vector, oil.vector)
cor.tibble = as.tibble(cor.p) 
cor.r
cor.tibble
```
So, reject the null hypothesis.
###Plot
```{r}
ggplot(data = cor.tibble)+
  geom_histogram(mapping= aes(x=value))+
  geom_vline(xintercept = cor.r, color="red") +
  labs(title = "Permutation Test for Oil and Arms")+
  xlab("Correlation")
```
### Map Function
```{r, warning=FALSE}
map.vector = oil[oil$country == "Iraq" |
                         oil$country == "Iran",]

map.vecotr = map(map.vector, mean)
map.vecotr$OP
```
### Contributions
Finished my individual section.

----------------------------------
