---
title: "lab 11 Feeling Dangerous"
author: "Fin,Nahom,Chun,Jake"
date: "11/7/2019"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(babynames)
babynames18 <- read_csv("2018babynames.txt", 
                        col_types = cols( Name = col_character(),
                                          Gender = col_character(),
                                          Count = col_double()))
total2018 <- sum(babynames18$Count)

babynames18 <- babynames18 %>% 
  mutate(prop = Count/total2018, year = 2018) %>%
  rename(n = Count, sex = Gender, name = Name) %>%
  select(year, sex, name, n, prop)
  
babynames <- full_join(babynames18, babynames)

```

# Individual plots of our name’s popularity over time

## Fin

## Jake
 
## Chun
Bob = babynames %>% filter(str_detect(name, "^Bob"))
qplot(x = year, y = n,data = Bob)
## Nahom

```{r}
Nahom_3 <- babynames%>%
  filter(str_detect(name, "^Nah"))
nahom_prop <- Nahom %>% group_by(year) %>% 
  summarize(total_prop = sum(prop))
Nahom_comb <- babynames %>% 
  filter(str_detect(name, "^Nah[ou]m$"))
nahom_cprop <- Nahom %>% group_by(year) %>% 
  summarize(total_prop = sum(prop))
  
  
  ggplot(nahom_prop)+
  geom_point(mapping= aes(year, total_prop))
  geom_point(data=nahom_prop, mapping = aes(x=year, y=total_prop))
  ```

#Ariel and Rachel regexs

## Fin

## Jake
```{r, include=FALSE}
rachel <- babynames %>% 
  filter(str_detect(name, "^R[aey]+ch[aey]+l[le]*$")) %>%
  filter(sex!='M') %>%
  filter(year %in% list(1973,1988,1990)) %>%
  group_by(year) %>%
  summarise(name_variants =n())

ariel <- babynames %>% 
  filter(str_detect(name, "^A[r]+[iy][ae]l[lae]*$")) %>%
  filter(sex!='M') %>%
  filter(year %in% list(1973,1988,1990)) %>%
  group_by(year) %>%
  summarise(name_variants =n())
```
In 1973 there were 13 variations of the name Rachel. In 1988 and 1990 there were 17 variations of the name Rachel. In 1973 there were 3 variations of the name Ariel.  In 1988 there were 11, and in 1990 there were 12 variations of the name Ariel.

## Chun

## Nahom

```{r, setup, echo=TRUE}
a_r<- babynames %>% 
  filter(str_detect(name, "^A[r]+[iy][ae]l[lae]*") | str_detect(name, "^R[aey]+ch[aey]+l[le]*$")) %>%
  filter(sex=='F') %>%
  filter(year %in% list(1973, 1988, 1990, 2018)) %>%
  mutate(frist_letter = str_sub(name, 1, 1))
 
a_r %>%
   group_by(frist_letter, year)%>%
  summarise(name_count=n())


a_r %>% group_by(year) %>%
  summarise(tot_prop = sum(prop))
  
```

# The Little Mermaid Effect

## Fin

## Jake
 
## Chun

## Nahom


# Our names

## Fin

## Jake
 
## Chun

## Nahom


## Our recommendation for a boy and girl name for the clients

## Fin

## Jake
 
## Chun

## Nahom
 
