---
title: "lab # Feeling Dangerous"
author: "Fin,Nahom,Chun,Jake"
date: "A/B/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)

```
# Team Section

### Overall Questions 

#### Does smoking increase the rate of premature births in pregnant women?

#### Does smoking in pregnant mother cause overall lower birth weights for newborns at every gestitational age?

### Why it Matters

### Team Plots

### Conclusions

### Recommendations

### Preliminary Question and Awnser


# Individual Section 

#### Jake

### Individual Questions

### Plots 

### Findings and Contributions


----------------------------------

#### Fin Carlson

### Individual Questions

### Plots 
```{r}
ggplot(data=babies)+
  geom_bar(mapping=aes(x=Premature), fill="blue")+
  facet_wrap(~smoke)

ggplot(data=babies)+
  geom_point(mapping=aes(y=bwtoz, x=gestation, color=smoke))+
  facet_grid(~smoke)


  

```


### Findings

### Contributions 

----------------------------------------------------------------

#### Nahom

### Individual Questions

### Plots

### Findings

### Contributions

-----------------------------------


#### Chun

### Individual Question

### Plots  

### Findings

### Contributions
 

