---
title: "lab # Feeling Dangerous"
author: "Fin,Nahom,Chun,Jake"
date: "A/B/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)

```
# Team Section

### Overall Questions 

#### Does smoking increase the rate of premature births in pregnant women?

#### Does smoking in pregnant mother cause overall lower birth weights for newborns at every gestitational age?

### Why it Matters

### Team Plots

### Conclusions

### Recommendations

### Preliminary Question and Awnser


# Individual Section 

#### Jake

### Individual Questions

### Plots 

### Findings and Contributions


----------------------------------

#### Fin Carlson

### Individual Questions

### Plots 
```{r}
ggplot(data=babies)+
  geom_bar(mapping=aes(x=Premature), fill="blue")+
  facet_wrap(~smoke)

ggplot(data=babies)+
  geom_point(mapping=aes(y=bwtoz, x=gestation, color=smoke))+
  facet_grid(~smoke)


  

```


### Findings

### Contributions 

----------------------------------------------------------------

#### Nahom Taye

### Individual Questions
Does the mother's race affect a newborn to be Premature?

### Plots
```{r, echo=FALSE, message=FALSE}
ggplot(babies, aes(mrace,inc))+
geom_boxplot(aes(mrace,inc))+
facet_wrap(~Premature)+
theme(legend.position = "top")+
ggtitle("Income vs Premature birth")
```

### Findings

While white women are the median for both of women that have exprienced Premature birth or have not, they have same income for the minimum and median income to those that either had premature birth or not, while there is a noticeable difference in the maximum income to those women that have exprienced premature birth and women that have not exprienced Prematuer birth have 1 more maximum point income differce from those that have exprineced Premature birth. This helps better understand that income affects Premature birth as women with high income can better her chances of prevent Premature birth.  


### Contributions

-----------------------------------


#### Chun

### Individual Question
Is mother's weights affects Premature when she smokes based on her race.
### Plots  
ggplot(babies)+
  geom_jitter(mapping = aes(x=Premature, y=mpregwt, alpha=smoke) )+
  facet_wrap(~mrace)+
  ggtitle("Mother's Weight VS. Premature if she smoke or not based on Mother's race ")
### Findings

### Contributions
 

