---
title: "lab # Feeling Dangerous"
author: "Fin,Nahom,Chun,Jake"
date: "A/B/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)
babies <- na.omit(babies)


```
# Team Section

### Overall Questions 

#### Does smoking increase the rate of premature births in pregnant women?

#### Does smoking in pregnant mother cause overall lower birth weights for newborns at every gestitational age?

### Why it Matters

### Team Plots

### Conclusions

### Recommendations

### Preliminary Question and Awnser


# Individual Section 

#### Jake

### Individual Questions

### Plots 

### Findings and Contributions


----------------------------------

#### Fin Carlson

### Individual Questions

### Plots 
```{r}
ggplot(data=babies)+
  geom_bar(mapping=aes(x=Premature), fill="blue")+
  facet_wrap(~smoke)

ggplot(data=babies)+
  geom_point(mapping=aes(y=bwtoz, x=gestation, color=smoke))+
  facet_grid(~smoke)


  

```


### Findings

### Contributions 

----------------------------------------------------------------

#### Nahom Taye

### Individual Questions
Does the mother's income affect a newborn to be Premature?

### Plots
```{r, echo=FALSE, message=FALSE}
ggplot(babies, aes(mrace,inc))+
geom_boxplot(aes(mrace,inc))+
facet_wrap(~Premature)+
theme(legend.position = "top")+
ggtitle("Income vs Premature birth")
```

### Findings

While white women are the median for both of women that have exprienced Premature birth or have not, they have same income for the minimum and median income of those that either had premature birth or not, while there is a noticeable difference in the maximum income to those women that have exprienced premature birth and women that have not exprienced Prematuer birth have 1 more maximum point income differce from those that have exprineced Premature birth. This helps better understand that income affects Premature birth as women with high income can better her chances of prevent Premature birth.  


### Contributions

-----------------------------------


#### Chun

### Individual Question
Is mother's race affects Premature when she smokes.
### Plots 
```{r} 
babies$mrace <- as.factor(babies$mrace)
babies$Premature <- as.factor(babies$Premature)
babies$smoke <- as.factor(babies$smoke)
ggplot(babies)+
  geom_bar(mapping = aes(x=Premature, fill=smoke), position = "fill" )+
  facet_grid(~mrace)+
  ggtitle(" Smoking VS. Premature based on Mother's race ")
### Findings
From the graph we can see that some white females'(from chart 2, 3 and 4) and  mixed races females' babies where born early mainly becaused they were still smoking when they were pregnant.  However, for example, in chart 1 we see that none of the mothers were smoking but still had premature babies.  Also from chart 0-4, we can see that lots of white mothers were still smoking when they were pregnant and from chart 5 - 8, most of white(or rest of the white), black, mexican ,  and asia mothers were not smoking( or never smoked before).  However, it is really hard to say if it is smoking that caused the premature babies.

### Contributions
 I created a chart with geom_bar to show the relation between smoker and premature with mother's races effected.  I use color to seperated if they are smoker or not.  Although most white women smoke and rest of the mothers from other races don't smoke but it is really hard to tell if smoking cause premature babies.

