---
title: "lab 3 Feeling Dangerous"
author: "Fin,Nahom,Chun,Jake"
date: "9/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
# Load and view the data
milk <- read.csv("/Users/fin/Desktop/tidytuesday-master/data/2019/2019-01-29/state_milk_production.csv")
head(milk)

milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)
```
# Team Section

### Overall Question 
##The year when the most milk was produced in the United States.
##The year when the least milk was produced in the United States.
```{r}
SumUS=milk%>% group_by(year) %>%summarise(Sum=sum(milk_million))%>% as_tibble()
NS1=arrange(SumUS, desc(Sum))
head(NS1, 1)
tail(NS1, 1)
```
# US produced most milk in 2017. US produced least milk in 1975
##In 2017, report the 5 states that produced the most milk.
##In 2017, report the 5 states that produced the least milk.
```{r}
milk2017 <- milk %>% filter(year == 2017)
SumState=milk2017%>% group_by(state) %>%summarise(milk_million)%>% as_tibble()
NS2=arrange(SumState, desc(milk_million))
head(NS2, 5)
tail(NS2, 5)
```

### Team Plot

```{r}
milkMI <- milk %>%
  filter(state == "Michigan")
ggplot(data = milkMI, aes(x = year, y = milk_million, color = state)) +
  geom_point() +
  geom_smooth()+
  ggtitle('Pounds of milk over time in MI') +
  xlab('Year') +
  ylab('Milk Produced (Million lb)')
```

### Conclusion



# Individual Section 

#### Jake

### Individual Question


### Plots 
```{r}

```

### Findings and Contributions


----------------------------------

#### Fin Carlson

### Individual Question



### Plots 

```{r echo=FALSE, message=FALSE }


```
 
```{r}

```
 
### Findings


### Contributions 

 

----------------------------------------------------------------

#### Nahom

### Individual Question



```{r echo=FALSE, message=FALSE}

```

### Plots

```{r}

```

### Findings



### Contributions



-----------------------------------


#### Chun


### Plots  

```{r}
milk1995 <- milk %>%
  filter(year == 1995)
```
1st
```{r}
  ggplot(data = milk1995, aes(x = milk_million)) +
  geom_histogram(bins = 20) + 
  ggtitle('Histogram of milk produced in 1995 by state')
```
2nd
```{r}
summary(milk1995)
```
3rd
```{r}
Summarize=milk %>% group_by(state) %>%
  summarise(Med = median(milk_million), Avg = mean(milk_million), Sum=sum(milk_million)) %>%
  as_tibble()
arrange(Summarize, desc(Sum))
```
### Findings
From the the plot for MI, we can see that milk produced rate increases significiently from the year 2000 and it is increasing linearly from year to year. Probrably because the rate of birth increased or more and more people moved into MI.
