---
title: "lab9_FeelingDangerous"
author: "Fin"
date: "10/24/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)

#loading and tidying data
data <- read.csv("nba_2017_nba_players_with_salary.csv")
data <- data %>% select(2:39)
data1 <- data %>% rename_at(vars(starts_with("X")), 
                                        funs(str_remove(., "X")))
data1 <-as_tibble(data1)


```
# Team Section

### Overall Question and Its Importance
Our teams objective is to determine the most overvalued and undervalued players for the 2017 NBA Regular Season. We will examine a number of box score statistics to better inform Front Office decisons for NBA franchises. This information will allow general managers to better understand the current value of their players for upcoming contract negoitation. Even more importantly, the information can be used when assesing both trade and draft talent.

### Dataset Description
The dataset being anaylzed contains the 2017-2018 season averages for each player in the NBA. In addition to traditional box scores, the data also includes many advanced statistics as well salary for the year. These two factors will be very important in our research and the dataset was picked with this in mind. All data comes from the NBA offical website however it has been scraped into a csv file for usability.

### Team Findings
It was determined

```{r}
```

### Conclusion and Recommendation

### Ethical Implications
 
# Individual Section 

#### Jake

### Individual Question

How do the salaries vary for the six players who average the highest points per game? In addition, what is the relationship between minutes played per game and averaged points for different positions across the NBA?


```{r, include=FALSE}
point10 <- data1 %>% filter(POINTS >= 10)

pos <- point10 %>% filter(POSITION=="C"  | POSITION == "PG" | POSITION == "SF")

Tseven <- data1 %>% filter(POINTS >=27)
```

### Plots
```{r, echo=FALSE}
p <- ggplot(data= Tseven, aes(x=PLAYER , y=SALARY_MILLIONS, fill=PLAYER))+
  geom_bar(stat="identity")+
  ggtitle('Salaries for Players with 26PPG or more in 2017')+
  xlab('Athletes')+
  ylab('Salary in Millions')+
  scale_fill_discrete(name="Delay") +
  theme(axis.text.x = element_text(angle = 30))
p + scale_fill_brewer(palette="Blues")

points <- ggplot(data=pos)+
  geom_point(mapping=aes(x=POINTS, y=MP, colour=POSITION))+
  geom_smooth(mapping=aes(x=POINTS, y=MP, colour=POSITION))
print(points + ggtitle("Points vs Minutes Players (averaging 10PPG+)"))

age <- ggplot(data=pos)+
  geom_point(mapping=aes(x=AGE, y=MP, colour=POSITION))+
  geom_smooth(mapping=aes(x=AGE, y=MP, colour=POSITION))
print(age + ggtitle("Age vs. Minutes Players (averaging 10PPG+)"))


```

### Findings and Contributions
To begin, I wanted to narrow the scope of my individual findings towards the best players is in the NBA; the players who are scoring 10 points or more per game. Then I filtered for different sizes of NBA player including the samllest, medium and the largest. To be more specific, point gaurd, small forward, and  center. I then ran minutes played aginst the age of those players and points per game in the 2017 season. I easily see the correlation I set color for position a geom to the position.

----------------------------------

#### Fin Carlson

### Individual Question
Which players are most undervalued as measured by their minutes played?

### Plots

```{r include=FALSE}
#EDA for Games Played vs. salary
ggplot(data=data1)+
  geom_point(mapping=aes(x=GP, y=SALARY_MILLIONS))+
  geom_smooth(mapping=aes(x=GP, y=SALARY_MILLIONS))
  
  ##data is to clumped to make assumptions, GP not a good factor
  summary(data1$GP)
```

```{r, include=FALSE}
#EDA for minutes per game vs. salary
ggplot(data=data1)+
  geom_point(mapping=aes(x=MPG, y=SALARY_MILLIONS))+
  geom_smooth(mapping=aes(x=MPG, y=SALARY_MILLIONS))

#code for anaylzing value based on mpg  
  
  ##avg salary 7.3
  ##avg mpg 21.5
  ##avg wins 2.86
  ##med wins 1.4
summary(data1$MPG)
summary(data1$SALARY_MILLIONS)
summary(data1$WINS_RPM)

  ##avg salary= 10.5 if mpg>21
min_above <-data1 %>% filter(MPG>= 21.57)

data1 %>% filter(MPG>= 21.57) %>% summarise(mean_data=mean(SALARY_MILLIONS))

  ##creates group of players who play more than average and are paid less than average
minutesSalary <-data1 %>% filter(MPG>= 21.57 & SALARY_MILLIONS<=10.5) 


  
  ##visualization of group
ggplot(data=minutesSalary)+
  geom_point(mapping=aes(x=MPG, y=SALARY_MILLIONS))+
  geom_smooth(mapping=aes(x=MPG, y=SALARY_MILLIONS))

  ##further filtering based on graph results

summary(minutesSalary$MPG) #avg mpg for this group is 27.8
summary(minutesSalary$SALARY_MILLIONS) #avg salary for this group is 4.85

#creating even smaller group based on group average
Finalminutes1<- minutesSalary %>% 
  filter(MPG>= 27.78 & SALARY_MILLIONS<=4.851) %>% arrange(MPG) 


##filters out truly worse than avg players using group avg of holistic stat 
Finalminutes2<- Finalminutes1 %>% filter(WINS_RPM>=2.86) 


##seeing who makes the most, and who is most benefitial 
leastPaid <- Finalminutes2 %>% 
  select(PLAYER, POSITION, MPG, SALARY_MILLIONS, WINS_RPM) %>% 
  arrange(SALARY_MILLIONS)
leastPaid

Best <- leastPaid %>% arrange(desc(WINS_RPM))
Best


## Further filtering undervalued group to show the top 3
min_above %>% filter(WINS_RPM>10)  #10/342 .03 percent of players have winshare>10 
Finalminutes3<- leastPaid %>% filter(WINS_RPM>10) 
Finalminutes3

```

```{r, echo=FALSE}
  ##visualization of undervalued group vs. total (mpg>21)
ggplot(min_above,aes(x=SALARY_MILLIONS,y=MPG))+
  geom_point() +
  geom_smooth() +
  geom_point(data=Finalminutes2,colour='red') 

    ##visualization of undervalued group with top 3 highlited (mpg>21)
ggplot(Finalminutes2,aes(x=SALARY_MILLIONS,y=MPG))+
  geom_point() +
  geom_point(Finalminutes3, mapping=aes(x=SALARY_MILLIONS,y=MPG, color=PLAYER))
   

#Under Valued Players are mostly younger.
ggplot(min_above,aes(x=AGE,y=MPG))+
  geom_point() +
  geom_smooth() +
  geom_point(data=Finalminutes2,colour='red') 

```

### Findings
My first graph highlights the group of players I determined as undervaluled as shown by the red clustering. I determined this by finding average levels for players and filtering by those who were above the average. While my analysis was focused on Salary and minutes played I also filtered by Wins by RPM as a means of assessing the talent of my smaller groups. I found 10 Wins by RPM to be the 97 percentile for the overall data and used this as my benchmark for the most undervalued of those seen in red. This is shown in my second graph and from this I determined Giannis Antentokounmpo, Nikola Jokic, and Rudy Gobert as the most undervalued players for the 2017 NBA season. My final graph shows that those in the undervalue group tend to be younger. This is likely due to the salary rules in the NBA that prevent large contracts for newer players.

### Contributions 

----------------------------------------------------------------

#### Nahom

### Individual Question

### Plots

```{r}

```

### Findings


### Contributions


-----------------------------------


#### Chun

### Individual Question
How to value the SG player?
#Since I don't really watch basketball game, so, from what my understanding, SG is a position to get all the points. To see if it is correct, I want to compare FG., 3P., 2P., FT. and POINTS between SG and overall players.
```{r}
SG=data1%>%filter(POSITION=='SG')%>%arrange(SALARY_MILLIONS)
summary(SG$FG.)#mean=0.4123
summary(SG$`3P.`)#mean=0.3531
summary(SG$`2P.`)#mean=0.4553
summary(SG$FT.)#mean=0.7896
summary(SG$POINTS)#mean=9.625

summary(data1$FG.)#mean=0.4461
summary(data1$`3P.`)#mean=0.3070
summary(data1$`2P.`)#mean=0.4867
summary(data1$FT.)#mean=0.7485
summary(data1$POINTS)#mean=9.446
```
#From the data above, we can see that SG should have a better 3-Point skill and ability to get points than the other players.

#Next, I want to see what effect the value of a SG player. I will acreat a plot to compare multiple variables.
```{r}
###Plots
ggplot(SG,aes(x=SALARY_MILLIONS,y=`3P.`))+
  geom_point() +
  geom_smooth()
ggplot(SG,aes(x=SALARY_MILLIONS,y=`3P`))+
  geom_point() +
  geom_smooth()
ggplot(SG,aes(x=SALARY_MILLIONS,y=`3PA`))+
  geom_point() +
  geom_smooth()

ggplot(SG,aes(x=SALARY_MILLIONS,y=AGE))+
  geom_point() +
  geom_smooth()
   

ggplot(SG,aes(x=SALARY_MILLIONS,y=POINTS))+
  geom_point() +
  geom_smooth()
ggplot(SG,aes(x=SALARY_MILLIONS,y=W))+
  geom_point() +
  geom_smooth()
```

#From the chart above, we can see that most of the variables and the salary will increases linearly together. But ones these variables like 3P., Wins 3PA, 3P reach their threshold, the relation will break. However, from POINTS vs SALARY_MILLIONS plot, we can see that SG who has higher points gets higher salary.
